---
title: "PM_final_project"
output: 
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#PART 1 - LOGISTIC REGRESSION AND NAIVE BAYES

```{r}
### Library Imports ###
library(glmnet)
library(caret) #for confusion matrix
library(e1071)
library(data.table)

### DATA LOADING , TEST/TRAIN SPLIT####

raw=read.csv("cleansed_data.csv") 
raw = na.omit(raw)

raw=raw[,-1] #removing index column

#Converting to factors
factor_vars1 = c(2, 3, 4, 24)
for (i in factor_vars1) {
  raw[[i]]<-as.factor(raw[[i]])
}

set.seed(1)
train = sample(1:nrow(raw),(0.8*nrow(raw)))
test = (-train)

train.data <- raw[train,]
test.data <- raw[test,]


### VARIABLE IMPORTANCE USING LASSO CV ###

x=model.matrix(def_pay∼. + SEX*MARRIAGE + SEX*EDUCATION + MARRIAGE,raw)[,-1] ## Adding interaction terms
y=raw$def_pay

var_names = names(x[1,])
var_count = rep(0, length(var_names))

for(i in c(1:10)){
  train = sample(1: nrow(x), (0.80)*nrow(x))
  test=(-train)
  cvfit.lasso = cv.glmnet(x[train,],y[train],
                          alpha=1, family = "binomial", standardize = TRUE)
  
  lasso_coef = predict(cvfit.lasso, newx=x[test,],
                       s = "lambda.min", type="coefficients")
  var_used_index = lasso_coef[-2]
  var_used_index = ifelse(var_used_index ==0, FALSE, TRUE)
  var_count[var_used_index] = var_count[var_used_index] +1
  print(paste0("done", i))
}

var_used_index_n = ifelse(var_count == 10, TRUE, FALSE)
names(data.frame(x))[var_used_index_n]

```


```{r}
### LOGISTIC REGRESSION ###
logit.model <- glm(def_pay~ LIMIT_BAL.+SEX+MARRIAGE+PAY_1+PAY_2+
                     PAY_3+PAY_4+PAY_5+PAY_6+BILL_AMT1.+PAY_AMT1.+PAY_AMT2.+PAY_AMT4.+PAY_AMT6.
                   ,data=train.data, family = binomial(link = "logit"))
logit.prob <- predict(logit.model,test.data,type = 'response')

threshold.logit <- 0.50
pred.logit <- ifelse(logit.prob > threshold.logit,1,0)

confusionMatrix(table(pred.logit, test.data$def_pay))

```


```{r}

### NAIVE BAYES MODEL ###
naive.model = naiveBayes(def_pay~LIMIT_BAL.+SEX+MARRIAGE+PAY_1+PAY_2+PAY_3+PAY_4+PAY_5+PAY_6+BILL_AMT1.
                         +PAY_AMT1.+PAY_AMT2.+PAY_AMT4.+PAY_AMT6., data=train.data, type="raw")

naive.prob <- predict(naive.model, test.data, type='raw')
naive.prob<-data.table((naive.prob))

threshold.naive <- 0.50
pred.naive <- ifelse(naive.prob$`1` > threshold.naive,1,0)

confusionMatrix(table(pred.naive, test.data$def_pay))

```


#PART 2 - KNN CLASSIFICATION

### KNN CLASSIFICATION WITHOUT NORMALIZATION

### Reading the data

```{r}
rm(list=ls())
library(tinytex)
credit_default1=read.csv('cleansed_data.csv',header=T)
credit_default1=credit_default1[-1]
names(credit_default1)
```

### Converting categorical variables 

```{r}
factor_vars1 = c(2, 3, 4, 24, c(6:11))
for (i in factor_vars1) {
credit_default1[[i]]<-as.factor(credit_default1[[i]])
}

```

### Splitting the data

```{r}
library(kknn)
library(class)
set.seed(33) 
tr1 = sample(c(1:dim(credit_default1)[1]), 20000)
train1 = credit_default1[tr1,]
test1 = credit_default1[-tr1,]
y_train1=credit_default1[tr1,24]
y_test1=credit_default1[-tr1,24]
```

### KNN using out of sample predictions for k=50

```{r}
set.seed(33)
library(caret)
library(precrec)
library(ROCit)
library(plotROC)
library(ggplot2)

knn1=knn(train1,test1,cl=y_train1,k=50,prob=FALSE,use.all=FALSE)
knn1_table=table(knn1,y_test1)
accuracy <- function(x){sum(diag(x)/(sum(rowSums(x)))) * 100}

x=accuracy(knn1_table)
precision=posPredValue(knn1,y_test1,positive='1')
recall=sensitivity(knn1,y_test1,positive = '1')
F1=(2*precision*recall)/(precision+recall)

print(knn1_table)
cat("Accuracy is",x,"and misclassification rate is",100-x)
cat("Precision is",precision,"and Recall is",recall)
cat("F1 score is",F1)
precrec_obj <- evalmod(scores = as.numeric(knn1), labels = y_test1)
autoplot(precrec_obj)
```

### Finding the best k value

```{r}
set.seed(33)
tr1 = sample(c(1:dim(credit_default1)[1]), 20000)
train1 = credit_default1[tr1,]
test1 = credit_default1[-tr1,]
y_train1=credit_default1[tr1,24]
y_test1=credit_default1[-tr1,24]
k_list1=c(50,100,150,200)
k_list1
y=NULL
for(i in k_list1){
  knn2 = knn(train1,test1,cl=y_train1,k=i)
  knn2_table=table(knn2,y_test1) 
  
  x=accuracy(knn2_table)
  precision=posPredValue(knn2,y_test1,positive='1')
  recall=sensitivity(knn2,y_test1,positive = '1')
  F1=(2*precision*recall)/(precision+recall)
  
  x1=100-x
  y=c(y,x1)
}
best = which.min(y)
plot(k_list1,y,type="b",xlab="K Value",col="blue",ylab="Misclassification rate",lwd=2,cex.lab=1.2)
cat("Best k value is",k_list1[best],"with misclassification rate of",y[best])
precrec_obj <- evalmod(scores = as.numeric(knn2), labels = y_test1)
autoplot(precrec_obj)
```

### K fold cross validation with kcv=10

```{r}
set.seed(33)
train1 = credit_default1
test1 = credit_default1
y_train1=credit_default1[tr1,24]
y_test1=credit_default1[-tr1,24]
n=dim(credit_default1)[1]
k_list1=c(20,50,100,150,200,250)
kcv = 10
n0 = round(n/kcv,0)
set=1:n
used = NULL
y1=matrix(0,kcv,6)
y2=matrix(0,kcv,6)
y3=matrix(0,kcv,6)
for(j in 1:kcv){
  if(n0<length(set)){val = sample(set,n0)}
  if(n0>=length(set)){val=set}
    train_i = train1[-val,]
    test_i = test1[val,]
    y_train_i=credit_default1[-val,24]
    y_test_i=credit_default1[val,24]
    for(i in 1:6){
     knn3 = knn(train_i,test_i,cl=y_train_i,k=k_list1[i])
     knn3_table=table(knn3,y_test_i)
     x1=accuracy(knn3_table)
     x2=100-x1
     precision=posPredValue(knn3,y_test_i,positive='1')
     recall=sensitivity(knn3,y_test_i,positive = '1')
     F1=(2*precision*recall)/(precision+recall)
     
     #cat(x2,"for k value",k_list1[i],"and fold",j,"\n")
     y1[j,i]=x2
     y2[j,i]=precision
     y3[j,i]=recall
    }
  used = union(used,val)
  set = (1:n)[-used]
  cat(j,'\n')
}  
my1=apply(y1,2,mean)
my2=apply(y2,2,mean)
my3=apply(y3,2,mean)
cat("Misclassification rate values:",my1)
cat("Precision values:",my2)
cat("Recall values:",my3)
best = which.min(my1)
plot(k_list1,my1,xlab="K value",ylab="Misclassification rate",col=4,lwd=2,type="l",cex.lab=1.2,main=paste("kfold(",kcv,")"))
cat("Best k value is",k_list1[best],"with misclassification rate of",my1[best])
```

### K fold cross validation with kcv=5

```{r}
set.seed(33)
train1 = credit_default1
test1 = credit_default1
y_train1=credit_default1[tr1,24]
y_test1=credit_default1[-tr1,24]
n=dim(credit_default1)[1]
k_list1=c(20,50,100,150,200,250)
kcv = 5
n0 = round(n/kcv,0)
set=1:n
used = NULL
y1=matrix(0,kcv,6)
y2=matrix(0,kcv,6)
y3=matrix(0,kcv,6)
for(j in 1:kcv){
  if(n0<length(set)){val = sample(set,n0)}
  if(n0>=length(set)){val=set}
    train_i = train1[-val,]
    test_i = test1[val,]
    y_train_i=credit_default1[-val,24]
    y_test_i=credit_default1[val,24]
    for(i in 1:6){
     knn4 = knn(train_i,test_i,cl=y_train_i,k=k_list1[i])
     knn4_table=table(knn4,y_test_i)
     x1=accuracy(knn4_table)
     x2=100-x1
     precision=posPredValue(knn4,y_test_i,positive='1')
     recall=sensitivity(knn4,y_test_i,positive = '1')
     F1=(2*precision*recall)/(precision+recall)
     
     #cat(x2,"for k value",k_list1[i],"and fold",j,"\n")
     y1[j,i]=x2
     y2[j,i]=precision
     y3[j,i]=recall
    }
  used = union(used,val)
  set = (1:n)[-used]
  cat(j,'\n')
}  
my1=apply(y1,2,mean)
my2=apply(y2,2,mean)
my3=apply(y3,2,mean)
cat("Misclassification rate values:",my1)
cat("Precision values:",my2)
cat("Recall values:",my3)
best = which.min(my1)
plot(k_list1,my1,xlab="K value",ylab="Misclassification rate",col=4,lwd=2,type="l",cex.lab=1.2,main=paste("kfold(",kcv,")"))
cat("Best k value is",k_list1[best],"with misclassification rate of",my1[best])
```

### KNN CLASSIFICATION WITH NORMALIZATION

### Reading the data

```{r}
rm(list=ls())
credit_default2=read.csv('cleansed_data.csv',header=T)
credit_default2=credit_default2[-1]

dim(credit_default2)
```

### Converting categorical variables 

```{r}
factor_vars2 = c(2, 3, 4, 24, c(6:11))
for (i in factor_vars2) {
credit_default2[[i]]<-as.factor(credit_default2[[i]])
}

```

### Normalization

```{r}
normalize <- function(x) {
return ((x - min(x)) / (max(x) - min(x))) }
normalize_vars=c(1,5,c(12:23))
for (i in normalize_vars) {
credit_default2[[i]]=normalize(credit_default2[[i]])
}
summary(credit_default2)
```

### Splitting the data

```{r}
library(kknn)
library(class)
set.seed(33) 
tr2 = sample(c(1:dim(credit_default2)[1]), 20000)
train2 = credit_default2[tr2,]
test2 = credit_default2[-tr2,]
y_train2=credit_default2[tr2,24]
y_test2=credit_default2[-tr2,24]
```

### KNN using out of sample predictions for k=50

```{r}
set.seed(33)
knn5=knn(train2,test2,cl=y_train2,k=50,prob=FALSE,use.all=FALSE)
knn5_table=table(knn5,y_test2)
accuracy <- function(x){sum(diag(x)/(sum(rowSums(x)))) * 100}

x=accuracy(knn5_table)
precision=posPredValue(knn5,y_test2,positive='1')
recall=sensitivity(knn5,y_test2,positive = '1')
F1=(2*precision*recall)/(precision+recall)

print(knn5_table)
cat("Accuracy is",x,"and misclassification rate is",100-x)
cat("Precision is",precision,"and Recall is",recall)
cat("F1 score is",F1)
precrec_obj <- evalmod(scores = as.numeric(knn5), labels = y_test2)
autoplot(precrec_obj)
```

### Finding the best k value

```{r}
set.seed(33)
tr2 = sample(c(1:dim(credit_default2)[1]), 20000)
train2 = credit_default2[tr2,]
test2 = credit_default2[-tr2,]
y_train2=credit_default2[tr2,24]
y_test2=credit_default2[-tr2,24]
k_list2=c(5,10,15,20,50,100)
k_list2
y=NULL
for(i in k_list2){
  knn6 = knn(train2,test2,cl=y_train2,k=i)
  knn6_table=table(knn6,y_test2) 
  x=accuracy(knn6_table)
  precision=posPredValue(knn6,y_test2,positive='1')
  recall=sensitivity(knn6,y_test2,positive = '1')
  F1=(2*precision*recall)/(precision+recall)
  
  x1=100-x
  y=c(y,x1)
}
best = which.min(y)
plot(k_list2,y,type="b",xlab="K Value",col="blue", ylab="Misclassification rate",lwd=2,cex.lab=1.2)
cat("Best k value is",k_list2[best],"with misclassification rate of",y[best])
precrec_obj <- evalmod(scores = as.numeric(knn6), labels = y_test2)
autoplot(precrec_obj)
```

### K fold cross validation with kcv=10

```{r}
set.seed(33)
train2 = credit_default2
test2 = credit_default2
y_train2=credit_default2[tr2,24]
y_test2=credit_default2[-tr2,24]
n=dim(credit_default2)[1]
k_list2=c(5,10,20,50,100,150)
kcv = 10
n0 = round(n/kcv,0)
set=1:n
used = NULL
y1=matrix(0,kcv,6)
y2=matrix(0,kcv,6)
y3=matrix(0,kcv,6)
for(j in 1:kcv){
  if(n0<length(set)){val = sample(set,n0)}
  if(n0>=length(set)){val=set}
    train_i = train2[-val,]
    test_i = test2[val,]
    y_train_i=credit_default2[-val,24]
    y_test_i=credit_default2[val,24]
    for(i in 1:6){
     knn7 = knn(train_i,test_i,cl=y_train_i,k=k_list2[i])
     knn7_table=table(knn7,y_test_i)
     
     x1=accuracy(knn7_table)
     x2=100-x1
     precision=posPredValue(knn7,y_test_i,positive='1')
     recall=sensitivity(knn7,y_test_i,positive = '1')
     F1=(2*precision*recall)/(precision+recall)
     
     cat(x2,"for k value",k_list2[i],"and fold",j,"\n")
     y1[j,i]=x2
     y2[j,i]=precision
     y3[j,i]=recall
    }
  used = union(used,val)
  set = (1:n)[-used]
  cat(j,'\n')
}  
my1=apply(y1,2,mean)
my2=apply(y2,2,mean)
my3=apply(y3,2,mean)
cat("Misclassification rate values:",my1)
cat("Precision values:",my2)
cat("Recall values:",my3)
best = which.min(my1)
plot(k_list2,my1,xlab="K value",ylab="Misclassification rate",col=4,lwd=2,type="l",cex.lab=1.2,main=paste("kfold(",kcv,")"))
cat("Best k value is",k_list2[best],"with misclassification rate of",my1[best])
```

### K fold cross validation with kcv=5

```{r}
set.seed(33)
train2 = credit_default2
test2 = credit_default2
y_train2=credit_default2[tr2,24]
y_test2=credit_default2[-tr2,24]
n=dim(credit_default2)[1]
k_list2=c(5,10,20,50,100,150)
kcv = 5
n0 = round(n/kcv,0)
set=1:n
used = NULL
y1=matrix(0,kcv,6)
y2=matrix(0,kcv,6)
y3=matrix(0,kcv,6)
for(j in 1:kcv){
  if(n0<length(set)){val = sample(set,n0)}
  if(n0>=length(set)){val=set}
    train_i = train2[-val,]
    test_i = test2[val,]
    y_train_i=credit_default2[-val,24]
    y_test_i=credit_default2[val,24]
    for(i in 1:6){
     knn7 = knn(train_i,test_i,cl=y_train_i,k=k_list2[i])
     knn7_table=table(knn7,y_test_i)
     
     x1=accuracy(knn7_table)
     x2=100-x1
     precision=posPredValue(knn7,y_test_i,positive='1')
     recall=sensitivity(knn7,y_test_i,positive = '1')
     F1=(2*precision*recall)/(precision+recall)
     
     cat(x2,"for k value",k_list2[i],"and fold",j,"\n")
     y1[j,i]=x2
     y2[j,i]=precision
     y3[j,i]=recall
    }
  used = union(used,val)
  set = (1:n)[-used]
  cat(j,'\n')
}  
my1=apply(y1,2,mean)
my2=apply(y2,2,mean)
my3=apply(y3,2,mean)
cat("Misclassification rate values:",my1)
cat("Precision values:",my2)
cat("Recall values:",my3)
best = which.min(my1)
plot(k_list2,my1,xlab="K value",ylab="Misclassification rate",col=4,lwd=2,type="l",cex.lab=1.2,main=paste("kfold(",kcv,")"))
cat("Best k value is",k_list2[best],"with misclassification rate of",my1[best])
```

### KNN with normalization and SMOTE

### Reading the data

```{r}
rm(list=ls())

credit_default3=read.csv('cleansed_data.csv',header=T)
credit_default3=credit_default3[-1]

```

### Converting categorical variables

```{r}
factor_vars3 = c(2, 3, 4, 24, c(6:11))
for (i in factor_vars3) {
credit_default3[[i]]<-as.factor(credit_default3[[i]])
}

```

### Normalization

```{r}
normalize <- function(x) {
return ((x - min(x)) / (max(x) - min(x))) }
normalize_vars=c(1,5,c(12:23))
for (i in normalize_vars) {
credit_default3[[i]]=normalize(credit_default3[[i]])
}

```

### Splitting the data

```{r}
library(kknn)
library(class)
set.seed(33) 
tr3 = sample(c(1:dim(credit_default3)[1]), 20000)
train3 = credit_default3[tr3,]
test3 = credit_default3[-tr3,]
y_train3=credit_default3[tr3,24]
y_test3=credit_default3[-tr3,24]
prop.table(table(y_test3))
```

```{r}
library(DMwR)
library(dplyr)
train3$def_pay <- as.factor(train3$def_pay)
train3 <- SMOTE(train3$def_pay ~ ., train3, perc.over = 100, perc.under=200)
train3$def_pay <- as.numeric(train3$def_pay)
train3$def_pay=ifelse(train3$def_pay==2,1,0)
prop.table(table(train3$def_pay))
```

```{r}
library(caret)
library(precrec)
library(ROCit)
library(plotROC)
library(ggplot2)

set.seed(33)
knn9=knn(train3,test3,cl=train3$def_pay,k=50,prob=FALSE,use.all=FALSE)
knn9_table=table(knn9,y_test3)
accuracy <- function(x){sum(diag(x)/(sum(rowSums(x)))) * 100}

x=accuracy(knn9_table)
precision=posPredValue(knn9,test3$def_pay,positive='1')
recall=sensitivity(knn9,test3$def_pay,positive = '1')
F1=(2*precision*recall)/(precision+recall)


print(knn9_table)
cat("Accuracy is",x,"and misclassification rate is",100-x)
cat("Precision is",precision,"and Recall is",recall)
cat("F1 score is",F1)
precrec_obj <- evalmod(scores = as.numeric(knn9), labels = y_test3)
autoplot(precrec_obj)
```

### K fold cross validation with kcv=10

```{r}
set.seed(33)
test3 = credit_default3
y_test3=credit_default3[-tr3,24]
n=dim(credit_default3)[1]
k_list3=c(5,10,20,50,100,150)
kcv = 10
n0 = round(n/kcv,0)
set=1:n
used = NULL
y1=matrix(0,kcv,6)
y2=matrix(0,kcv,6)
y3=matrix(0,kcv,6)
for(j in 1:kcv){
  if(n0<length(set)){val = sample(set,n0)}
  if(n0>=length(set)){val=set}
    train_i = train3[-val,]
    test_i = test3[val,]
    y_test_i=credit_default3[val,24]
    for(i in 1:6){
     knn10 = knn(train_i,test_i,cl=train_i$def_pay,k=k_list3[i])
     knn10_table=table(knn10,y_test_i)
     
     x1=accuracy(knn10_table)
     x2=100-x1
     precision=posPredValue(knn10,y_test_i,positive='1')
     recall=sensitivity(knn10,y_test_i,positive = '1')
     F1=(2*precision*recall)/(precision+recall)
     
     cat(x2,"for k value",k_list3[i],"and fold",j,"\n")
     y1[j,i]=x2
     y2[j,i]=precision
     y3[j,i]=recall
    }
  used = union(used,val)
  set = (1:n)[-used]
  cat(j,'\n')
}  
my1=apply(y1,2,mean)
my2=apply(y2,2,mean)
my3=apply(y3,2,mean)
cat("Misclassification rate values:",my1)
cat("Precision values:",my2)
cat("Recall values:",my3)
best = which.min(my1)
plot(k_list3,my1,xlab="K value",ylab="Misclassification rate",col=4,lwd=2,type="l",cex.lab=1.2,main=paste("kfold(",kcv,")"))
cat("Best k value is",k_list3[best],"with misclassification rate of",my1[best])
```

#PART - 3 DECISION TREES

```{r}
# Fitting Classification Trees
rm(list=ls())
library(tree)
df = read.csv('cleansed_data.csv')
df$ID = NULL

train_sample = sample(c(1:dim(df)[1]), 20000)
train = df[train_sample,]
test = df[-train_sample,]
ytrain = train$def_pay
ytest = test$def_pay

tree.credit = tree(factor(def_pay)~.,data=train)
summary(tree.credit)
plot(tree.credit)
text(tree.credit, pretty =0)


```
```{r}
#In order to properly evaluate the performance of a classification tree on these data
#We must estimate the test error rather than simply computing the training error.
tree.pred = predict(tree.credit, test ,type ="class")
table(tree.pred, ytest)

```
```{r}
#The function cv.tree() performs cross-validation in order to 
#cv.tree() determine the optimal level of tree complexity;
set.seed (3)
cv.credit =cv.tree(tree.credit,FUN=prune.misclass)

par(mfrow =c(1,2))
plot(cv.credit$size, cv.credit$dev ,type="b")
plot(cv.credit$k, cv.credit$dev ,type="b")

```
```{r}
prune.credit =prune.misclass (tree.credit, best =2)
plot(prune.credit)
text(prune.credit,pretty =0)
summary(prune.credit)

```
```{r}

tree.pred.prune=predict (prune.credit, test ,type="class")
table(tree.pred.prune, ytest)


```

```{r}
################################################################################
library(tree)
library(rpart)
library(rpart.plot)

#fit a big tree using rpart.control
big.tree = rpart(def_pay ~ .,method="class",data=train,
                 control=rpart.control(minsplit=5,cp=.0005))
nbig = length(unique(big.tree$where))
cat('size of big tree: ',nbig,'\n')
rpart.plot(big.tree)

#--------------------------------------------------
#look at cross-validation
par(mfrow=c(1,1))
plotcp(big.tree, col = 'orange')

#--------------------------------------------------
#show fit from some trees
bestcp=big.tree$cptable[which.min(big.tree$cptable[,"xerror"]),"CP"]
cat('bestcp: ',bestcp,'\n')

ptree = prune(big.tree,cp=bestcp)
rpart.plot(ptree)
```

# PART 4 - RANDOM FOREST

```{r}
library(randomForest)

# Read in data
data = read.csv("cleansed_data.csv")

# Take Id out of dataset
credit.card = data[, !names(data) == "ID"]

# Factorize the categorical variables
cate_vari = c(2:4,6:11,24)
for (i in cate_vari){
  credit.card[[i]] = as.factor(credit.card[[i]])
}

# Split the data into training set and testing set
# We agree on using 2/3 data for train
train = sample(1:nrow(credit.card), 20000)
test = (-train)
```

```{r}
# Pick n_tree using cross validation.
# We tried: n.list = seq(50,1000,length=20) | k=6 for extual result which shows in graph and yieds best result ntree = 600. But it was time consuming. So for this turn-in code just test 3 btree value.
n.list = c(30, 60, 120)
k = 3
n.samp = round(nrow(credit.card)/k) # determine the train size for k-fold
iAccuracyMx = matrix(0, k, length(n.list)) # matrix records in sample accuracy
accuracyMx = matrix(0, k, length(n.list)) # out-of-sample accuracy matrix 
used = NULL
set = 1:nrow(credit.card) 
for (ki in 1:k){
  # determine number of each fold in case data cant be split evenly for each fold
  if (length(set)>=n.samp) {val=n.samp}  
  if (length(set)<n.samp) {val=length(set)}
  test.cv = sample(set, val)
  train.cv = (-test.cv)
  for (ni in 1:length(n.list)){
    cat('Now is processing:', ki, ni, "\n")
    set.seed(7)
    rf.fit = randomForest(def_pay~., data = credit.card[train.cv,], 
                           ntree = n.list[ni], importance = T, mtry=23)
    rf.pred = predict(rf.fit, newdata = credit.card[test.cv,])
    # Geting in-sample accuracy
    iAccuracyMx[ki, ni] = mean(rf.fit$predicted==credit.card$def_pay[train.cv])
    # Getting out-of sample accuracy
    accuracyMx[ki, ni] = mean(rf.pred==credit.card$def_pay[test.cv]) 
    
  used = union(used, test.cv) # update valication set for other folds 
  set = (1:nrow(credit.card))[-used]
  }
}
```

```{r}
# Calculate the average error for each ntree
mError = 1-apply(accuracyMx, 2, mean) # Geting out-of-sample misclassification
imError = 1-apply(iAccuracyMx, 2, mean) # Geting in sample misclassification
intree = which.min(mError) # index for best ntree

# Plot the both in sample and out of sample error rates
plot(n.list, mError, col="red", pch=17, type = 'b', 
     main="Cross Valication for Tunning ntree",
     xlab = "Number of Trees (B)",
     ylab = "Error Rate")
lines(n.list, imError, col="blue", pch=16, type = 'b', lty=2)
points(n.list[intree], min(mError), col = "green", pch = 17, cex = 1.8)
legend("topright", 
       c("Out-of-sample error", "In-sample error", "Best ntree: 120"),
       col = c("red", "blue", "green"), lty = 1:2)
```

```{r}
# Selecting best mtry using cross validation
mtry.list = c(2,3,10) # tried mtry.list = 1:23 | k =5 for actual analysis & got best mtry = 3. Here using simple version to demenstrate, saving running time. 
k2 = 3
n.samp2 = round(nrow(credit.card)/k2)
iAccuracyMx2 = matrix(0, k2, length(mtry.list))
accuracyMx2 = matrix(0, k2, length(mtry.list))
used = NULL
set = 1:nrow(credit.card)
for (ki in 1:k2){
  if (length(set)>=n.samp2) {val=n.samp2}
  if (length(set)<n.samp2) {val=length(set)}
  test.cv = sample(set, val)
  train.cv = (-test.cv)
  for (mti in 1:length(mtry.list)){
    cat('Now is processing:', ki, mti, "\n")
    set.seed(7)
    rf.fit = randomForest(def_pay~., data = credit.card[train.cv,], 
                           ntree = 120, importance = T, 
                           mtry = mtry.list[mti]) # in actual analysis use ntree=600
    rf.pred = predict(rf.fit, newdata = credit.card[test.cv,])
    iAccuracyMx2[ki, mti]= mean(rf.fit$predicted==credit.card$def_pay[train.cv])
    accuracyMx2[ki, mti] = mean(rf.pred==credit.card$def_pay[test.cv])
  used = union(used, test.cv)
  set = (1:nrow(credit.card))[-used]
  }
}
```

```{r}
# Calculate the average error for each mtry
mError2 = 1-apply(accuracyMx2, 2, mean)
imError2 = 1-apply(iAccuracyMx2, 2, mean)
imtry = which.min(mError2) # index of best mtry

# Plot the both in sample and out of sample errors
plot(mtry.list, mError2, col="red", pch=17, type = 'b', 
     main="Cross Valication for Tunning mtry",
     xlab = "Number of Variable Used for Each Split",
     ylab = "Error Rate")
lines(mtry.list, imError2, col="blue", pch=16, type = 'b', lty=2)
points(mtry.list[imtry], min(mError2), col = "green", pch = 17, cex = 1.8)
legend("topright", 
       c("Out of sample error", "In sample error", "Best mtry: 3"),
       col = c("red", "blue", "green"), lty = 1:2)
```

```{r}
# Use the best mtry = 3, and best ntree = 600 retrain the data
rf.fit_best = randomForest(def_pay~., data = credit.card[train,],mtry=3,ntree = 120, importance = T) # in analysis we use ntree=600
# Predict test set 
set.seed(7)
rf.pred_best = predict(rf.fit_best, newdata = credit.card[test,])
table(rf.pred_best, credit.card$def_pay[test]) # make confusion matrix
```

```{r}
cat("Random forest correct rate with mtry = 3:", # Showing accuracy
    mean(rf.pred_best ==credit.card$def_pay[test]))
```

```{r}
# Check importance
importance(rf.fit_best)
```

```{r}
varImpPlot(rf.fit_best)
```

#PART 5 - XG BOOST

### process data

```{r data}
library(xgboost)
df = read.csv('cleansed_data.csv')
df$ID = NULL


# factorize
# factor_vars1 = c(2, 3, 4, 24, c(6:11))
# for (i in factor_vars1) {
# df[[i]]<-as.factor(df[[i]])
# }

## sampling
set.seed(33)
tr = sample(c(1:dim(df)[1]), 20000)
train = df[tr,]
test = df[-tr,]
y_tr = train$def_pay
y_ts = test$def_pay

dim(df)
summary(df)
```

### xgb model before tuning

```{r xgb}
xgb <- xgboost(data = data.matrix(train[,1:23]), 
 label = data.matrix(y_tr), 
 eta = 1,
 max_depth = 6, 
 nround=300, 
 subsample = 1,
 colsample_bytree = 1,
 lambda = 1,
 seed = 33,
 eval_metric = "error",
 objective = "binary:logistic",
)
```

### result before tuning on test dataset, output is accuracy rate

```{r predict}
y_test <- predict(xgb, data.matrix(test[,1:23]))
y_test[y_test < 0.5] = 0
y_test[y_test > 0.5] = 1
1 - sum(y_test == y_ts)/length(y_ts)
```

```{r}
y_prediction = y_test
tbl = table(y_prediction, y_ts)
print(tbl)
print(paste("Precision: ", tbl[2,2]/sum(tbl[2,])))
print(paste("Recall: ", tbl[2,2]/sum(tbl[,2])))

library(caret)
precision <- posPredValue(as.factor(y_prediction), as.factor(y_ts), positive="1")
recall <- sensitivity(as.factor(y_prediction), as.factor(y_ts), positive="1")
print(paste("Precision: ", precision))
print(paste("Recall: ", recall))

```

### xgb model after tuning

```{r xgb2}
xgb <- xgboost(data = data.matrix(train[,1:23]), 
 label = data.matrix(y_tr), 
 eta = 0.05,
 max_depth = 3, 
 nround=10, 
 subsample = 1,
 colsample_bytree = 1,
 lambda = 10,
 seed = 33,
 eval_metric = "error",
 objective = "binary:logistic",
)
```

### result after tuning on test dataset, output is accuracy rate

```{r predict2}
y_test <- predict(xgb, data.matrix(test[,1:23]))
y_test[y_test < 0.5] = 0
y_test[y_test > 0.5] = 1
1 - sum(y_test == y_ts)/length(y_ts)

```

```{r}
y_prediction = y_test
tbl = table(y_prediction, y_ts)
print(tbl)
print(paste("Precision: ", tbl[2,2]/sum(tbl[2,])))
print(paste("Recall: ", tbl[2,2]/sum(tbl[,2])))

precision <- posPredValue(as.factor(y_prediction), as.factor(y_ts), positive="1")
recall <- sensitivity(as.factor(y_prediction), as.factor(y_ts), positive="1")
print(paste("Precision: ", precision))
print(paste("Recall: ", recall))
```



### feature importance

**Gain** 

is the improvement in accuracy brought by a feature to the branches it is on. The idea is that before adding a new split on a feature X to the branch there was some wrongly classified elements, after adding the split on this feature, there are two new branches, and each of these branch is more accurate (one branch saying if your observation is on this branch then it should be classified as 1, and the other branch saying the exact opposite).

**Cover**

measures the relative quantity of observations concerned by a feature.

**Frequency**

is a simpler way to measure the Gain. It just counts the number of times a feature is used in all generated trees. You should not use it (unless you know why you want to use it).

```{r feature importance}

xgb.importance(model = xgb)

```

### CV: iterates by nrounds

```{r CV}
nrounds = 500

credit = xgb.DMatrix(as.matrix(train[,1:23]), label = as.matrix(y_tr))

cv <- xgb.cv(data = credit, nrounds = nrounds, nfold = 5, metrics = "error",
                  max_depth = 6, eta = 0.3, objective = "binary:logistic",
                subsample = 1, colsample_bytree = 1, seed = 33, 
                lambda = 20, verbose = 0)
print(cv)
```

### visualize CV result: we don't really need a lot of trees

Conclusion: 10 trees would work ok

```{r CV2}
plot(c(1:nrounds), cv[4]$evaluation_log$test_error_mean, xlab = 'number of trees', ylab = 'error', type = 'l')
```

### how does the depth of each tree affects error

Conclusion: 3 splits for each tree seems to be optimized

```{r depthoftrees}
start_time <- Sys.time()
minError = c()

for(i in c(2:10)){
        cv <- xgb.cv(data = credit, nrounds = 10, nfold = 5, metrics = "error",
                  max_depth = i, eta = 0.3, objective = "binary:logistic",
                subsample = 1, colsample_bytree = 1, seed = 33, 
                lambda = 1, verbose = 0)
        minError = c(minError, min(cv[4]$evaluation_log$test_error_mean))        
}

plot(c(2:10), minError, xlab = 'depth of trees', ylab = 'error', type = 'l')
end_time <- Sys.time()
end_time - start_time

```

### how lambda affects error

conclusion: lambda 10 would work ok

```{r lambda}
start_time <- Sys.time()
testVariable = 'lambda'
testRange = seq(1, 100, 10)
minError = c()

for(i in testRange){
        cv <- xgb.cv(data = credit, nrounds = 10, nfold = 5, metrics = "error",
                  max_depth = 3, eta = 0.3, objective = "binary:logistic",
                subsample = 1, colsample_bytree = 1, seed = 33, 
                lambda = i, verbose = 0)
        minError = c(minError, min(cv[4]$evaluation_log$test_error_mean))        
}

plot(testRange, minError, xlab = testVariable, ylab = 'error', type = 'l')

end_time <- Sys.time()
end_time - start_time
```

### how eta affects error

```{r eta}
start_time <- Sys.time()
testVariable = 'eta'
testRange = seq(0.0, 1, 0.1)
minError = c()

for(i in testRange){
        cv <- xgb.cv(data = credit, nrounds = 10, nfold = 5, metrics = "error",
                  max_depth = 3, eta = i, objective = "binary:logistic",
                subsample = 0.8, colsample_bytree = 0.8, seed = 33, 
                lambda = 10, verbose = 0)
        minError = c(minError, min(cv[4]$evaluation_log$test_error_mean))        
}

plot(testRange, minError, xlab = testVariable, ylab = 'error', type = 'l')

end_time <- Sys.time()
end_time - start_time
```